A Framework for Equivalence of Cast Calculi and a new Cast Representation

### 2020-10-05 paper.tex

Fig.~\ref{fig:R-machine-syntax} defines machine 
states $s$, including a definition of values $v$ and value typing.
Fig.~\ref{fig:R-machine-dynamics} defines the transition relation of the CEK
machine (\CMachine{B}).
We name the machine after \textbf{c}ast calculus.
\todo[inline]{JS: explain the name choice of $\mathcal{R}$.}
\todo[inline]{KC: I want to space-efficient machine to be S, so ... B might be a better name 
because it reminds people of the $\lambda B$ calculus.
Or $T$ for type-based cast?
Or $C$ for cast-calculus?
Or $D$ for definitional semantic?
How do you think?}
\todo[inline]{JS: B or C both sound like good choices for the reasons
you give.}
\todo[inline]{KC: Let's go with C to avoid confusion with Blame strategy.}
%
This machine is parameterized over blame strategy $B$, which ranges over
Lazy D (Fig.~\ref{fig:blame-strategy-LazyD}) and Lazy UD 
(Fig.~\ref{fig:blame-strategy-LazyUD}).
\todo[inline]{JS: I don't think $B$ is restricted to just those two.
Can't it be any blame strategy that satisfies certain conditions?}
\todo[inline]{KC: Technically a $B$ should define a predicate over pretypes, which specifies
what is injectable, and a apply-cast function. Of course, the set of values that the apply-cast
function works with depends on the set of injectable types.
\url{https://github.com/Gradual-Typing/Equivalence-of-Cast-Calculi/blob/master/equivalence-of-cast-calculi/R/BlameStrategies.agda}
Do you want me to put these contraints in the paper?}
\todo[inline]{JS: Yes.}
%
The transitions involving casts are highlighted in red and described in
more detail below. The other transitions are standard for a CEK
machine for an extended simply typed lambda calculus~\citep{Felleisen:2009aa}.
%
Recall that CEK machines involve two kinds of transitions: those that
dive further into an expression (looking for a redex) and push an
entry onto the continuation and those that return a value to the
current continuation and possibly perform a computation.  That is, the
machine state is either in an evaluating $\sexpr{e}{E}{K}$ or
returning $\scont{v}{K}$ configuration, respectively. Additionally,
there are two halting configurations: \shalt{v} halts with a value,
while \error{l} halts with a blame label.  Observables include all the
value constructors and blame labels, as in the work of
\citet{siek2012interpretations}.  The function that converts values to
observables ($observe(v) = o$) is defined in the obvious way.

@@

% All cast representation shall be proved correct, that is, besides reducing 
% space consumption, using a cast representation should not affect the results of 
% program execution.
In this section we present a framework for proving the 
correctness of cast representations. The framework includes an abstract data 
type named Cast ADT, which specifies the operators of cast representations, 
a CEK machine $\SMachine{C}$ parameterized over a Cast ADT,
which uses the cast operators implemented by $C$ to work with casts, 
and theorems relating $\SMachine{C}$ to $\CMachine{B}$.
%
Subsection~\ref{sec:cast-adt-and-space-efficient-cek} describes Cast ADT and the 
machine \SMachine{C}. We conjecture that the machine \SMachine{C} is 
space-efficient provided that the cast representation $C$ performs compression.
%
Subsection~\ref{sec:good-cast-adt} defines two subset of Cast ADT, namely Lazy D 
Cast ADT and Lazy UD Cast ADT. As their names suggest, if $C$ is a Lazy D Cast 
ADT, executing a program in $\SMachine{C}$ would give the same result as in 
$\LDMachine$. Similarly, if $C$ is a Lazy UD Cast ADT, executing a program in 
$\SMachine{C}$ would give the same result as in $\LUDMachine$.
%
We prove the theorems relating \SMachine{C} and \CMachine{B} in
Subsection~\ref{sec:S-correct}.
The framework is written in the Agda language~\citep{norell2008dependently}.

>>>>

Fig.~\ref{fig:R-machine-syntax} defines machine 
states $s$, including a definition of values $v$ and value typing.
Fig.~\ref{fig:R-machine-dynamics} defines the transition relation of the CEK
machine (\CMachine{B}).
We name the machine after \textbf{c}ast calculus.
%
This machine is parameterized over blame strategy $B$, which ranges over
Lazy D (Fig.~\ref{fig:blame-strategy-LazyD}) and Lazy UD 
(Fig.~\ref{fig:blame-strategy-LazyUD}).
\todo[inline]{JS: I don't think $B$ is restricted to just those two.
Can't it be any blame strategy that satisfies certain conditions?}
\todo[inline]{KC: Technically a $B$ should define a predicate over pretypes, which specifies
what is injectable, and a apply-cast function. Of course, the set of values that the apply-cast
function works with depends on the set of injectable types.
\url{https://github.com/Gradual-Typing/Equivalence-of-Cast-Calculi/blob/master/equivalence-of-cast-calculi/R/BlameStrategies.agda}
Do you want me to put these contraints in the paper?}
\todo[inline]{JS: Yes.}
%
The transitions involving casts are highlighted in red and described in
more detail below. The other transitions are standard for a CEK
machine for an extended simply typed lambda calculus~\citep{Felleisen:2009aa}.
%
Recall that CEK machines involve two kinds of transitions: those that
dive further into an expression (looking for a redex) and push an
entry onto the continuation and those that return a value to the
current continuation and possibly perform a computation.  That is, the
machine state is either in an evaluating $\sexpr{e}{E}{K}$ or
returning $\scont{v}{K}$ configuration, respectively. Additionally,
there are two halting configurations: \shalt{v} halts with a value,
while \error{l} halts with a blame label.  Observables include all the
value constructors and blame labels, as in the work of
\citet{siek2012interpretations}.  The function that converts values to
observables ($observe(v) = o$) is defined in the obvious way.

@@

% All cast representation shall be proved correct, that is, besides reducing 
% space consumption, using a cast representation should not affect the results of 
% program execution.
In this section we present a framework for proving the 
correctness of space-efficient cast representations.
The framework includes an abstract data 
type named Cast ADT, which specifies the operators of cast representations, 
a CEK machine $\SMachine{C}$ parameterized over a Cast ADT,
which uses the cast operators implemented by $C$ to work with casts, 
and theorems relating $\SMachine{C}$ to $\CMachine{B}$.
%
Subsection~\ref{sec:cast-adt-and-space-efficient-cek} describes Cast ADT and the 
machine \SMachine{C}. We conjecture that the machine \SMachine{C} is 
space-efficient provided that the cast representation $C$ performs compression.
%
Subsection~\ref{sec:good-cast-adt} defines two subset of Cast ADT, namely Lazy D 
Cast ADT and Lazy UD Cast ADT. As their names suggest, if $C$ is a Lazy D Cast 
ADT, executing a program in $\SMachine{C}$ would give the same result as in 
$\LDMachine$. Similarly, if $C$ is a Lazy UD Cast ADT, executing a program in 
$\SMachine{C}$ would give the same result as in $\LUDMachine$.
%
We prove the theorems relating \SMachine{C} and \CMachine{B} in
Subsection~\ref{sec:S-correct}.
The framework is written in the Agda language~\citep{norell2008dependently}.
